\section{Introduction}
\label{sec:intro}

%With the rising popularity of cloud computing, the popularity of
%virtualization is steadily growing as a way to implement
%\emph{multi-tenancy}.

In recent years, with the rising popularity of cloud computing,
virtualization became crucially important as a way to provide
\emph{multi-tenancy}. While \emph{Infrastructure as a Service} (IaaS)
model assumes provisioning an entire operating system using virtual
machine monitors (\eg Google Compute Engine), \emph{Platform as a
Service} (PaaS) model typically relies on a language runtimes (\eg
Google App Engine). The~advantage of the former is a finer degree of
control over the entire application stack while the latter allows for
better resource utilization. A~modern hypervisor uses hardware
virtualization technology to enforce isolation and resource allocation;
in case of language runtime, ad-hoc sandboxing mechanism is typically
being used to provide the same guarantees, which results in a limited
number of supported platforms (\eg Java, Python and Go in case of Google
App Engine).

%As cloud providers can use multi-tenancy run large number of
%applications---potentially coming from different customers---under a
%single operating system instance. The existing PaaS platforms such as
%Google App Engine already uses multi-tenancy where several applications
%share the same physical or virtual machine.  However, to ensure the
%security and isolation, the number of supported platforms is limited to
%Java, Python and Go using custom sandboxed runtimes.

%The existing PaaS platforms such as Google App Engine already uses
%multi-tenancy where several applications share the same physical or
%virtual machine.  However, to ensure the security and isolation, the
%number of supported platforms is limited to Java, Python and Go using
%custom sandboxed runtimes.

In this position paper, we propose an alternative which combines
advantages of both approaches. We argue for an application virtual
machine capable of executing native code providing the standard POSIX
interface. We aim to combine \emph{software fault isolation}
(SFI)~\cite{wahbe:sosp93} for security and isolation properties with
microkernel-style service oriented architecture to provide process
management, file and network access. Based on top of Native
Client~\cite{yee:ieee-sp09}, state of the art industrial strength SFI
implementation, this solution can provide security and performance
assurances comparable to existing system virtual machines at a
fractional \emph{trusted code base} (TCB) size which allows for easier
verification and maintenance. 

Such execution environment could form a basis for a new kind of
POSIX compliant PaaS native platform, which will allow for execution of
high-performance applications. Such platform will be more efficient than
traditional virtualization and some language based sandboxes while more
flexible than Linux-based systems, akin to virtual machines.
Furthermore, native code support will enable use of existing C/C++
libraries and legacy code as well as extended instruction sets (\eg SSE
or AVX) for high-performance applications. Example of such service is
Google
Exacycle\footnote{See \url{http://googleresearch.blogspot.com/2011/04/1-billion-core-hours-of-computational.html} and
\url{http://research.google.com/university/exacycle_program.html}.},
which provides PaaS platform for large-scale, embarrasingly parallel
batch computations.

Furthermore, native code support together with full POSIX interface
allows running existing runtimes on top of our execution environment
without the necessity of ad-hoc sandboxing mechanism since the
environment itself already provides sandboxing capabilities. This has a
number of advantages.  First, it is possible to use ``vanilla'' language
runtimes which allows for upgrading to a newer version immediately after
they are released significantly reducing the risk of zero-day
vulnerabilities. Second, it allows provisioning of wider range of
runtimes potentially attracting more customers, even those using less
popular languages and platforms.  Third, it will allow use of native
extensions supported by most language runtimes, but not always available in
PaaS environments for security reasons.

%The solution based on Native Client would allow executing native
%POSIX-compliant code at near native speed while providing the same
%security and isolation guarantees. This will allow the use of wide
%range of existing legacy software without the overhead of running (and
%managing) the entire VM.

%i like NaCl as a posix-like platform-as-a-service a little better,
%though restricting to the posix-y subset sort of works wrt running on a
%real posix-y OS too, i guess.

%more efficient than virtualization and some language based sandboxes.
%more flexible than LBS, akin to VMs
%not just novelty, but TCB minimization

The rest of this paper is organized as follows.
Section~\ref{sec:overview} gives an overview of our approach, \ldots.
Then, Section~\ref{sec:prototype} presents \ldots. Finally,
Section~\ref{sec:related} discusses related work and
Section~\ref{sec:conclusion} concludes.

%Furthermore, our approach gives greater level control which is so
%important in cloud deployments.

%In recent years, we have witnessed a boom of virtualization. Most of the
%server infrastructure which is now being used runs virtualized.
%However, virtualization still failed to attract desktop market despite
%obvious advantages such as portability, security, reliability and
%isolation. We believe this is because the existing virtualization
%solutions are too heavyweight and cumbersome. They require the user run
%an entire operating system which is suitable for infrastructure
%provisioning in data centers as well as certain cloud applications, but
%is simply too cumbersome for desktop and client applications.

%In this paper, we propose a more lightweight alternative. Instead of
%running the full operating system, our goal is to combine aspects of
%microkernel and monolithic kernel design with software fault isolation
%techniques to provide lightweight solution for virtualization of
%existing applications in the user space. Furthemore, our proposed
%approach allows applications to be run on top of three major operating
%system without recompilation.
