% rubber: bibtex.crossrefs 10000
\documentclass[twocolumn,10pt]{article}

%
% Standard packages.
%
\usepackage{xspace}
\usepackage{color}
\usepackage{times}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage[hyphens]{url}
\usepackage{multirow}
\usepackage{flushend}
\usepackage{textcomp}

\usepackage{listings}
\lstset{
  language=C,
  captionpos=b,
  basicstyle=\ttfamily,
  commentstyle=\textit,
  keywordstyle=\bfseries,
  numbers=left,
  numberstyle=\tiny,
  numbersep=5pt,
  escapeinside={/*@}{@*/},
  tabsize=2,
  breaklines=true,
  extendedchars=true,
  columns=fullflexible,
  xleftmargin=16pt,
  showspaces=false,
  showtabs=false,
  showstringspaces=false,
}

%
% Standard margin stuff.  US letter.
%
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{\oddsidemargin}
\setlength{\footskip}{0.5in}
\setlength{\textheight}{8.5in}% 11in - 2x1in margins - .5 for page number
\setlength{\textwidth}{6.5in} % 8.5in - 2x1in margins

\newcommand{\textstt}[1]{{\small \texttt{#1}}\xspace}
\newcommand{\textssc}[1]{{\small \textsc{#1}}\xspace}

\newcommand{\todo}[1]{\noindent{\color{red}[{\bf todo:} #1]}}
\newcommand{\sidenote}[1]{{\marginpar{\bf\color{red}[#1]}}}
\newcommand{\gap}[2]{\uline{#1} \noindent{\color{red}[#2]}}

\newenvironment{structure}{\color{blue}
\begin{list}{\labelitemi}{\leftmargin=1em}{%
\setlength{\topsep}{0pt}\setlength{\itemsep}{0pt}\setlength{\parsep}{0pt}}}
{\end{list}}

\newenvironment{structure*}{\color{blue}
\begin{enumerate}\setlength{\topsep}{0pt}\setlength{\itemsep}{0pt}\setlength{\parsep}{0pt}}
{\end{enumerate}}

\newcommand{\ie}{\emph{i.e.},\ }
\newcommand{\eg}{\emph{e.g.},\ }
\newcommand{\etal}{\emph{et al.}\ }
\newcommand{\vs}{\emph{vs.}\ }
\newcommand{\etc}{\emph{etc.}\ }

\begin{document}

\title{\bf Microkernel in the User Space}
\author{
Petr Hosek\footnote{Work done while an intern at \textit{Google Inc}.}\\\textit{Imperial College London}
\and Bennet Yee\\\textit{Google Inc.}
}
\date{}
\maketitle
\thispagestyle{empty}

\begin{abstract}
In recent years, together with the rising popularity of cloud computing
we have witnessed a boom of virtualization. Most of the software
infrastructure which is being used nowadays runs virtualized in some
form. However, existing virtualization solutions are still too limited
despite obvious advantages such as portability, security and isolation.
Virtual machine monitors require the user to run an entire operating
system which is suitable for infrastructure provisioning in data centers
as well as certain cloud applications, but is simply too cumbersome for
desktop and client applications.  Language runtimes on the other hand
are typically targeting only a single high-level language (or family) making
them unsuitable for native applications.
 
In this paper, we propose an alternative approach. We aim to combine
aspects of microkernel and hybrid kernel design with software fault
isolation techniques to provide lightweight solution for virtualization
of existing POSIX based applications entirely in the user space.
Furthermore, our proposed approach provides portability without
significant performance overhead, allowing applications to be run on top
of three major operating systems without modifications or even
recompilation. \todo{Shall we say any operating system---even though
Native Client only supports Windows, Linux and OS X at the moment, it
should be possible to add support for other operating systems if
needed.  I'd rather not: there are OSes to which porting NaCl would be
difficult, e.g., missing some basic IPC mechanism.  Also, the
portability currently has an ISA-dependency in NaCl (vs lack of assembly in
PNaCl)---this is probably too much detail for the abstract.}
\end{abstract}

\input{intro}
\input{overview}
\input{prototype}
\input{related}
\input{conclusion}

{
  \bibliographystyle{abbrv}
  \bibliography{macros,references,crossreferences}
}

\end{document}
